---
title: "03-data-manipulation-NU"
author: "Nolan Ung"
date: "9/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Indexing and subsetting data frames

Our survey data frame has rows and columns (it has 2 dimensions), if we want to
extract some specific data from it, we need to specify the "coordinates" we
want from it. Row numbers come first, followed by column numbers. However, note
that different ways of specifying these coordinates lead to results with
different classes.

```{r load}
# install and load tidyverse
install.packages("tidyverse")
libary(tidyverse)

## load our data frame 
luciferase <- read_csv(file = "luciferase_toy_data.csv", col_types = "cdffdcdcffl")
print(luciferase)

#view data in Viewer 
View(luciferase)

```



```{r subsetting}
# first element in the first column of the data frame (as a vector)
luciferase[1, 1]   
# first element in the 6th column (as a vector)
luciferase[1, 6]   
# first column of the data frame (as a vector)
luciferase[, 1]    
# first column of the data frame (as a data.frame)
luciferase[1]      
# first three elements in the 7th column (as a vector)
luciferase[1:3, 7] 
# the 3rd row of the data frame (as a data.frame)
luciferase[3, ]    

```

### Subsetting by calling column names directly

Column names can be used directly to inpect a whole column one at a time. The column name is referenced using the `$`.  if we wanted to only see the enitre `Condition` Column we could use `csv_luciferase$Condition`

```{r}
csv_luciferase$Condition

```

```{r}
 #1. Create a `data.frame` (`luciferase_20`) containing only the data in row 20 of the `csv_luciferase` dataset.

# 2. Notice how `nrow()` gave you the number of rows in a `data.frame`?

      # Use that number to pull out just that last row in the data frame.
      # Compare that with what you see as the last row using `tail()` to make
      # sure it's meeting expectations.
      # Pull out that last row using `nrow()` instead of the row number.
      # Create a new data frame (`luciferase_last`) from that last row.



 # 3. Use `nrow()` to extract the row that is in the middle of the data
    #frame. Store the content of this row in an object named `luciferase_middle`.

# 4. Combine `nrow()` with the `:` notation above to reproduce the behavior offset
#    `tail(luciferase)`, keeping just the 40th through last rows of the luciferase
#    dataset.



```

## Factors

Factors represent categorical data. They are stored as integers associated with labels and they can be ordered or unordered. While factors look (and often behave) like character vectors, they are actually treated as integer vectors by R. So you need to be very careful when treating them as strings.

Once created, factors can only contain a pre-defined set of values, known as levels. By default, R always sorts levels in alphabetical order. For instance, if you have a factor with 2 levels:

```{r}
plate <- csv_luciferase$Plate_ID
print(plate)

```

R will assign 1 to the level "BIC_luc" and 2 to the level "ENZ_luc" (because B comes before E). You can see this by using the function levels() and you can find the number of levels using nlevels():

```{r}
# use the functions level() and nlevels() to examine the Plate_ID column

```

Sometimes, the order of the factors does not matter, other times you might want to specify the order because it is meaningful (e.g., “low”, “medium”, “high”), it improves your visualization, or it is required by a particular type of analysis. Here, one way to reorder our levels in the plate vector would be:
```{r}
levels(plate) # check current order

plate  <- factor(plate, levels = c('ENZ_luc', 'BIC_luc')) # reoder

levels(plate) # check after reordering
```

## viewing and Renaming factors

When your data is stored as a factor, you can use the plot() function to get a quick glance at the number of observations represented by each factor level. 

```{r}
plot(plate)

# let's rename them to remove the "_luc" since its redundant (labels)



# plot to make sure its correct



```

## Dplyr and Tidyr for  data manipulation

Dplyr and Tidyr were made for data manipulation and analysis and are the most common packages used for these tasks. 

lets start with our `luciferase` dataset. and we are going to be introduced to the common functions of dplyr as we manipulate our datset to get to the final product. 

Harish got his data of the luminometer and made it  into a csv and loaded it into R (like we previously did)  and now he has a data frame but he needs to clean up certain things and get some information out of it. 

when he was running the experiment he realized that something went wrong with the first replicate so he decided that we need to get rid of that. 

lets "filter" out replicate 1 using the function `filter()`

```{r}  


?dplyr::filter  # look up the filter function

filtered_luc <- filter(luciferase, !Rep == 1) #filter out rep 1 and asign it  to filtered_luc

 # view filtered_luc

 # summary or unique just to check


```

Let's do the same task, only now we are going to use pipes. the pipes symbol is `%>%`. If you use RStudio, you can type the pipe with <kbd>Ctrl</kbd> + <kbd>Shift</kbd> + <kbd>M</kbd> if you have a PC or <kbd>Cmd</kbd> + <kbd>Shift</kbd> + <kbd>M</kbd> if you have a Mac. a pipe takes the output of the function on the left and "pipes" it into the first argument of the function on the right. thus you can easily chain multiple functions  together. for example:

```{r pipe-example, echo=TRUE}

filtered_luc <- filter(luciferase, !Rep == 1)
summary_luc <- summary(filtered_luc)
print(summary_luc)


# the two lines above is the same as the line below 
summary_luc_pipe <- luciferase %>% filter(!Rep == 1) %>% summary()
print(summary_luc_pipe)

```

*NOTE:* pipes work better with some functions than others. All of the functions in dplyr and some in the other tidyverse packages all have the "data" argument as the first argument. For other packages outside of the tidyverse this isn't always the case. 

We notice the the TF booster is False in all  cases and isn't really  informative. lets remove that column. 
we can do this by using the `select()` function.  lets pipe out output from `filter()` to `select()`. 

```{r select, echo  = TRUE}
# use pipes,  filter and select to remove rep one and the TF_booster column. 

select_luc  <- luciferase 


```


lets say that we want to  show our `Signal` data as relative to the lowest value in `Signal`. we can use mutate to do this on the whole column. we also notice that the concentration units are different and that we would like to make them uniform. we can do both of these tasks using the function `mutate()`. mutate taeks a column does sometihing to it ad outputs another column. you can add this column to your data set or you can replace a column with it. 


```{r mutate}

# add a new column make signal relative to the lowest signal.  mutate(rel_signal = Signal / min(Signal))

mutate_min_luc <- 

# use mutate to multiply D_Concentration by 1000  - mutate(D_Concentration = D_Concentration * 1000)

mutate_luc <- 

  
# change the units to nM
mutate_luc$L_Units <- "nM"
mutate_luc$D_Units <- "nM"


```


