---
title: "05-visualizing-data"
author: "Nolan Ung"
date: "9/26/2019"
output: html_document
---

If not still in the workspace, load the data we saved in the previous lesson.


```{r load-data, eval=FALSE, purl=FALSE}
luciferase <- read_csv(file = "luciferase_toy_data.csv", col_types = "cdffdcdcffl")

```

## Plotting with **`ggplot2`**

**`ggplot2`** is a plotting package that makes it simple to create complex plots
from data in a data frame. We only need minimal changes if the underlying data change
or if we decide to change from a bar plot to a scatter plot. This helps in creating
publication quality plots with minimal amounts of adjustments and tweaking.

**`ggplot2`** functions like data in the 'long' format, i.e., a column for every dimension,
and a row for every observation. Well-structured data will save you lots of time
when making figures with **`ggplot2`**

ggplot graphics are built step by step by adding new elements. Adding layers in
this fashion allows for extensive flexibility and customization of plots.

To build a ggplot, we will use the following basic template that can be used for different types of plots:

```
 ggplot(data = <DATA>, mapping = aes(<MAPPINGS>)) +  <GEOM_FUNCTION>()
```

- use the `ggplot()` function and bind the plot to a specific data frame using the
      `data` argument

```{r, eval=FALSE, purl=FALSE}
ggplot(data = luciferase)
```

- define a mapping (using the aesthetic (`aes`) function), by selecting the variables to be plotted and specifying how to present them in the graph, e.g. as x/y positions or characteristics such as size, shape, color, etc.

lets say that we want to check the replicability of this sample. lets plot Signal vs. Rep

```{r, eval=FALSE, purl=FALSE}
ggplot(data = luciferase, mapping = aes(x = Rep, y = Signal))
```

- add 'geoms' – graphical representations of the data in the plot (points,
  lines, bars). **`ggplot2`** offers many different geoms; we will use some 
  common ones today, including:
  
      * `geom_point()` for scatter plots, dot plots, etc.
      * `geom_boxplot()` for, well, boxplots!
      * `geom_line()` for trend lines, time series, etc.  


To add a geom to the plot use the `+` operator. Because we have two continuous variables,
let's use `geom_point()` first:

```{r point}
ggplot(data = luciferase, mapping = aes(x = Rep, y = Signal)) + geom_point()

```

let check thecconsistency of each condition by coloring the dots by condition. we can do that by setting the `color` argument in the aes() to Condition. remeber: this works because COndition is a factor. any of these groupoing variables needs to be factor in order to work properly. 


```{r}

ggplot(data = luciferase, mapping = aes(x = Rep, y = Signal, color = Condition)) + geom_point()


```

If the data points overlap too much you can change the opacity with the `alpha` argument. Also plotting the y on a logrithmic scale will help strch out all the small values

```{r}

ggplot(data = luciferase, mapping = aes(x = Rep, y = Signal, color = Condition)) + 
  geom_point(alpha = .5) + 
  scale_y_log10()

```

Next, lets look at the distribution of Signal of each condition as a bar chart

```{r box_plot}

ggplot(data = luciferase, mapping = aes(x = Rep)) + 
  geom_bar(stat = 'count') 

```


```{r}
ggplot(data = luciferase, mapping = aes(x = Condition, y = Signal)) + 
  geom_boxplot() + 
  scale_y_log10() 

```

You can layer different geoms to create your own kind of plots. 

```{r}
ggplot(data = luciferase, mapping = aes(x = Condition, y = Signal)) + 
  geom_boxplot() + 
  scale_y_log10() + 
  geom_jitter()
```



**`ggplot2`** has a special technique called *faceting* that allows the user to split one
plot into multiple plots based on a factor included in the dataset. 

```{r facet}
ggplot(data = luciferase, mapping = aes(x = Condition, y = Signal)) + 
  geom_boxplot() + 
  scale_y_log10() + 
  geom_jitter() +
  facet_wrap(facets = vars(Plate_ID))
```

```{r facet}
ggplot(data = luciferase, mapping = aes(x = Condition, y = Signal, col = Exp_Date)) + 
  geom_boxplot() + 
  scale_y_log10() + 
  geom_jitter() +
  facet_wrap(facets = vars(Plate_ID))
```


## Make a heatmap with the pheatmap function

```{r}
# install and load the pheatmap package

# install.packages("pheatmap")
library(pheatmap)

rna_mat <- read_csv("Processed_data/rna_matrix.csv") %>% column_to_rownames(var = "gene_names")
pheatmap(rna_mat, scale = "row", cluster_rows = TRUE, show_rownames = FALSE)


```

```{r}

rna_df <- rna_mat %>% rownames_to_column('gene')
ggplot(data = rna_df, mapping = aes(x = `Res-1`, y = `Res-2`, label = gene)) + geom_text() 


```

